The Astropy Project: \\ Sustaining and Growing a Community-oriented Open-source Project and the Latest Major Release (v5.0) of the Core Package

### 2022-04-26 src/ms.tex
explain Astropy’s idiosyncratic mixin definition
Signed-off-by: nstarman <nstarkman@protonmail.com>

Within \astropypkg and \python, there are numerous ways to represent and store
array-valued data. Some of the differences are historical: \astropytable and
\astropyapidoc{io.fits.BinTableHDU}{io.fits.BinTableHDU} are not the same since
\astropyfits was originally developed in a separate package (\package{PyFITS})
yet both are for holding tabular data. Some differences are computational:
\package{numpy}, \package{cupy}, and \package{dask} arrays have almost identical
APIs, but are optimized for different use cases. Lastly, some differences are
inherent: \astropyQuantity, \astropyTime, and \astropySkyCoord\ represent
fundamentally different types of objects.

\astropytable is \astropypkg's most general container for array-valued data.
Rather than requiring component data be converted to a
\astropyapidoc{table.Column}{table.Column}, a well-defined protocol has been
developed for any array-valued data to be used in a \astropytable, without
having to convert the data. The protocol for ``mixing'' column types provides
the same familiar table API, yet loses none the object's data or attributes.
With this protocol, it is now possible to store \astropy native objects --
including \astropyTime, \astropyQuantity, and \astropySkyCoord\ -- within a
\astropyTable\ and write these to various file formats, such as FITS. For
objects not already covered by the ``mixing'' protocol, functions can be
registered with \astropyTable\ to enable the interoperability. As an example,
\astropypkg arelady implements the ``mixing'' functions to integrate
\package{dask} arrays with \astropyTable, allowing cloud-stored or cluster-scale
data to be used as a column in a \astropyTable.

>>>>

Within \astropypkg and \python, there are numerous different ways to represent
and store array-valued data. Some differences are historical: \astropytable and
\astropyapidoc{io.fits.BinTableHDU}{io.fits.BinTableHDU} are both for holding
tabular data but the latter was developed for and remains largely specific to
FITS. Some differences are computational: \package{numpy}, \package{cupy}, and
\package{dask} arrays have almost identical APIs, but are optimized for
different use cases. Lastly, some differences are inherent: \astropyQuantity,
\astropyTime, and \astropySkyCoord\ represent fundamentally different types of
objects.

\astropytable is \astropypkg's most general container for array-valued data.
Rather than requiring component data be converted to a
\astropyapidoc{table.Column}{table.Column}, a well-defined protocol has been
developed for any array-valued data to be used in a \astropytable, without
having to convert the data. The protocol for mixing column types, which astropy
terms ``mixin''-columns, provides the same familiar table API, yet loses none
the object's data or attributes. Note ``mixin'' here does not refer to Python's
inheritance definition. Rather, it is a composition protocol that makes it
possible to store \astropy native objects -- including \astropyTime,
\astropyQuantity, and \astropySkyCoord\ -- within a \astropyTable\ and write
these to various file formats, such as FITS. For objects not already covered by
the astropy mixin-protocol, functions can be registered with \astropyTable\ to
enable interoperability. As an example, \astropypkg already implements these
mixin functions to integrate \package{dask} arrays with \astropyTable, allowing
cloud-stored or cluster-scale data to be used as a column in a \astropyTable.
